---
title: "Progress Memo"
subtitle: |
  | Final Project 
  | Foundations of Data Science with R (STAT 359)
author: "Kayla Terrelonge"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---

::: {.callout-tip icon=false}

## Github Repo Link

To link to your github **repo**sitory, appropriately edit the example link below. Meaning replace `https://your-github-repo-url` with your github repo url. Suggest verifying the link works before submitting.

[https://github.com/STAT359-2024SU/359-final-project-kaylaterrelonge](https://github.com/STAT359-2024SU/359-final-project-kaylaterrelonge)

:::


## Data source

The data for my project will be webscrapped from basketball references and ESPN. The data will be a combination of three datasets, coaching data from Basketball Reference, Player data from ESPN, and championship data from Basketball Reference. These data sets will be joined by the unique key of team and season. The data for players includes statistics for the player performance averaged by season. Only players who appeared in 70% of games are included. The coaching data is a list of head coaches and which team the coached for a given season. The championship data contains information about who won the championship, the MVP of the championship and who the top performers were in the playoffs leading up to the championship.  ^[Data Sources --- 
[ESPN Data](https://www.espn.com/nba/stats/player/_/season/2021/seasontype/2)
[Basketball Reference](https://www.basketball-reference.com/playoffs/)]

## Prediction Problem

The prediction problem I have chosen to explore for this project is predicting the position of a player, based on their performance statistics. This can be beneficial for coaching and strategy staff because it can provide insight, especially when drafting/recruiting players. Players may be comfortable in one position, but their game stats may paint a better picture as to where they will thrive on the court. This is a classification problem, as `pos` is a factor variable with 6 levels.

## Data quality & complexity check

```{r}
#| label: load-data
#| message: false
library(tidyverse)
# players
nba_files <- dir("data/raw_data_players/", pattern = "\\.csv$", full.names = TRUE)
nba_files

nba_files_imported <-list()

for(i in seq_along(nba_files)){
  nba_files_imported[[i]] <- read_csv(nba_files[[i]])
}
players <- bind_rows(nba_files_imported)

# coaches
coaching_files <- dir("data/raw_data_coaches/", pattern = "\\.csv$", full.names = TRUE)
coaching_files

coaching_files_imported <-list()

for(i in seq_along(coaching_files)){
  coaching_files_imported[[i]] <- read_csv(coaching_files[[i]])
}

nba_coaching_data <- bind_rows(coaching_files_imported)

skimr::skim_without_charts(players)
```

The players dataset has 24 variables and 9600 observations and the coaching data has 755 observations and 30 variables

## Target variable analysis

```{r}
#| label: target-var
ggplot(players, mapping = aes(x = POS)) +
  geom_bar()
```
 
The target variable seems to be imbalanced for certain levels. It is heavily imbalanced for the level of GF. However, because of its low count, it may be beneficial to filter out the small group of players who are GF (guards/fowards).

## Potential data issues

When thinking about potential issues, the issue surrounding the imbalance of certain positions may arise. It can be difficult to fix this, as certain positions are rather specialized and dependent on other factors like player measurements. To handle this, I will use stratified sampling during the model building process to ensure that the imbalance is reflected in the training process.